---
title: "Web Scraping with rvest and RSelenium in R, Use Case"
format: html
editor: visual
---

## Loading libraries

```{r libraries,warning=FALSE,message=FALSE}
# Load packages
library(RSelenium)
library(rvest)
library(netstat)
library(tidyverse)
```

## Read the web page

```{r}
# Navigate to the webpage
# Using rvest, navigate to the web page
presidential_result_page <-
  read_html('https://www.stears.co/elections/2023/president/')

```

### 2023 Presidential results

-   Summarized results

```{r}

# extract all state names
summary_div <- presidential_result_page |> html_elements("div.mb-4")

summarised_result_tbl <- tibble(
candidate_name = summary_div |> 
  html_elements("article.border-gray-10") |> 
  html_elements("p.font-jakarta.text-p_15_15.font-medium.text-gray-800") |> 
  html_text2(),
# party
candidate_party = summary_div |> 
  html_elements("article.border-gray-10") |> 
  html_elements("div") |> 
  html_elements("p.font-jakarta.font-jakarta.text-p_15_15.font-light.text-gray-600") |> 
  html_text2(),
# total result
total_result = summary_div |> 
  html_elements("article.border-gray-10") |> 
  html_elements("div") |> 
  html_elements("p.font-jakarta.font-jakarta.text-p_12_12.font-light.text-gray-800") |> 
  html_text2(),
# percentage of result
pct_result = summary_div |> 
  html_elements("article.border-gray-10") |> 
  html_elements("div") |> 
  html_elements("p.font-jakarta.font-jakarta.text-p_12_12.font-light.text-gray-600.ml-1") |> 
  html_text2()
)

```

## Start Selenium server, and using the Chrome browser

```{r message=FALSE,warning=FALSE}
# remote driver session using rsDriver() from  the RSelenium package
rs_driver_object <- rsDriver(browser = 'chrome',
                             chromever = "118.0.5993.70",
                             # from netstat package
                             port = free_port(),
                             verbose = FALSE)

remDr <- rs_driver_object$client


# Navigate to the web page

election_result_url <- 'https://www.stears.co/elections/2023/president/'
remDr$navigate(election_result_url)

```

## Scrape all results down to State and LGA levels

```{r}

# find and click the "Select State" drop-down
state_dropdown <-
  remDr$findElement(using = 'id',
                    'headlessui-menu-button-:R356t6:')

# the dropdown is clicked to open it
state_dropdown$clickElement()

```

Get the page source, and the state web page URLs
```{r}
# scrape all state url from their generated anchors
page_source <- remDr$getPageSource()
page <- read_html(page_source[[1]])
state_anchor_hrefs <- page  |>  
  html_elements(".flex.flex-col.justify-start.align-middle") |> 
  html_elements("a.text-gray-800")  |> 
  html_attr("href")

print(state_anchor_hrefs)
```

Next, we concatenate the main URL to each element of the state_anchors_href vector to generate the state results full URLs

```{r}

state_results_urls <- paste("https://www.stears.co", state_anchor_hrefs, sep = "")

print(state_results_urls)
```

For example, first state Abia State

```{r}

abia_state_page <- read_html(state_results_urls[[1]])

# extract all state names
abia_result_tables <- abia_state_page |> 
  html_element("div.px-6.py-10.bg-white.md\\:px-8.md\\:py-6") |> 
  html_elements("section.mb-4") |> 
  html_table()

abia_lga_names <- abia_state_page |> 
  html_element("div.px-6.py-10.bg-white.md\\:px-8.md\\:py-6") |> 
  html_elements("section.mb-4") |> 
  html_elements("p.font-jakarta.font-jakarta.text-p_20.font-medium.-md\\:hidden.mt-10.capitalize") |> 
  html_text2()

# print(abia_lga_names)

abia_state_name <- abia_state_page |>  
  html_element("h1.w-full.h-12.font-bold.font-jakarta.text-\\[28px\\].leading-\\[40px\\]") |> 
  html_text2()

# print(abia_state_name)

# first lga i.e. Umuahia North
Umuahia_North_lga_tbl <- abia_result_tables[[1]] |> add_column(LGA=abia_lga_names[[1]],
                                             STATE=abia_state_name,
                             .before = "Candidate")

# print(Umuahia_North_lga_tbl)

```

## Get all results tables: states and lgas

```{r}
all_states_results <- list()
# loop through the url for each state
for (state_results_url in state_results_urls) {
  state_page <- read_html(state_results_url)
  
  # extract all state results by lga. The result for each lga is kept in a table
  # on the page
  result_tables <- state_page |> 
    html_element("div.px-6.py-10.bg-white.md\\:px-8.md\\:py-6") |> 
    html_elements("section.mb-4") |> 
    html_table()
  
  lga_names <- state_page |> 
    html_element("div.px-6.py-10.bg-white.md\\:px-8.md\\:py-6") |> 
    html_elements("section.mb-4") |> 
    html_elements("p.font-jakarta.font-jakarta.text-p_20.font-medium.-md\\:hidden.mt-10.capitalize") |> 
    html_text2()
  
  state_name <- state_page |>  
    html_element("h1.w-full.h-12.font-bold.font-jakarta.text-\\[28px\\].leading-\\[40px\\]") |> 
    html_text2()

  
  # Iterate through the list of tables and add the STATE and LGA columns
  for (i in 1:length(lga_names)) {
    # Reference each table using their indices
    table_index <- i
    current_table <- result_tables[[table_index]]
    
    # Add LGA and STATE columns to the current table using the vector element
    current_table["LGA"] <- lga_names[i]
    current_table["STATE"] <- state_name
    
    # Update the current table in the list
    result_tables[[table_index]] <- current_table
  }
  
  all_states_results <- append(all_states_results, result_tables)
  
}

print(all_states_results)
```

### Print out the result

```{r}
# Print the modified list of tables
print(all_states_results[[1]])
print(all_states_results[[100]])
print(all_states_results[[774]])

```

### Data manipulation

```{r}
# Bind all tables in the all_states_results together
combined_tibble <- bind_rows(all_states_results) |> 
  select(-`Votes Margin`) |> 
  replace_na(list(Votes="0",Percent="0%"))

print(combined_tibble)
 
```


